---
title: "DATA 550 Final Project"
author: "Siddhi Desai"
date: "`r Sys.Date()`"
output: html_document
---

```{r, location-data-load, include = FALSE}
here::i_am("Final_Report.Rmd")
```


# Introduction

The National Health and Nutrition Examination Survey (NHANES) is an ongoing effort by the CDC National Center for Health Statistics to collect cross-sectional data on the demographics, health, and nutrition characteristics of people living in the U.S. The NHANES collects demographic, dietary, examination, lab, and questionnaire data. The dataset in this report is a subset of the 2017 to March 2020 Pre-Pandemic NHANES data wherein specific variables regarding demographic variables and questionnaire answers on cancer were extracted. See \*Cleaning the Data\* for more information on the specific variables.

The analysis objectives were to (1) create a table summarizing the extracted demographic variables (age, race, gender, and income-to-poverty ratio (PIR)) stratified by whether or not cancer was present, and (2) create a bar graph of the percentage of people with cancer within each race/ethnicity.


## Tabular Analysis

### Table 1

```{r, table-1, message=FALSE, echo = FALSE}
table1 <- readRDS(
  file = here::here("output/table1.rds"))

table1
```

Table 1 (above) summarizes demographic variables for `r nrow(data)` people from the NHANES 2017 to March 2020 Pre-Pandemic dataset. The overall and cancer-stratified proportions of males and females were fairly equal (48% and 52%, respectively). The median age of those with cancer or malignancies was 70 years (IQR: 59, 79) while the median age of those without cancer or malignancies was 49 years (IQR: 34, 63). The family income to poverty ratio had a median(IQR) of 2.63(1.48, 4.97) among those with cancer or malignancies and 2.16(1.17, 4,14) among those without. When stratified by race/ethnicity, Non-Hispanic Whites made up the greatest percentage of individuals with cancer or malignancies (61%) followed by Non-Hispanic Blacks (18%), Mexican Americans (6.0%), Other Hispanics (5.9%), Non-Hispanic Asians (4.8%), and Other Races (4.7%).

## Graphical Analysis

### Figure 1

```{r, fig-1, message=FALSE, echo = FALSE}
#| fig.align = "center",
#| out.width = "600px"
knitr::include_graphics(
  here::here("output/figure1.png"))
```

Figure 1 (above) presents the percentage of people within each race/ethnicity category that had cancer or malignancies. The percentage was highest among Non-Hispanic Whites (18.5%) followed by Other Races (10.7%), Non-Hispanic Blacks (7.6%), Other Hispanics (6.7%), Mexican Americans (6.1%), and Non-Hispanic Asians (4.6%).
